Tensor denoising and completion based on ordinal observations

### 2020-02-19 old_files/old_preprint/miaoyan/ordinalT_supp_v2.tex
clean up

\subsection{Clustering result on HCP}
We perform $K$-means method to cluster brain nodes based on the suggested above method. We have the estimated latent parameter $\hat\Theta$ with the rank $(23,23,8)$ which minimizes the BIC value. The number of clusters is chosen to be eight from the elbow method in figure ~\ref{figure:elbow}.
The table in \ref{table:clustering} shows eight cluster groups and their entries. Most of brain nodes fall into the cluster $\#1$ and the cluster $\#2$ which can be represented as the left side brain and the right side brain. From the group $\#3$ to the group $\#8$ , entries of each cluster share the same name encoded in the data and play a similar role in the brain. For example, the group $\#3$ and the group $\#4$ are the left and the right side of the regions called Supramarginal gyrus. Those regions are known to be in charge of interpreting tactile sensory data and involved in perception of space and limbs location \cite{carlson2012physiology,reed1994nature}. The figure \ref{figure:brain image} shows the nodes in the dataset on the brain image. Each node is connected by edges within the same cluster and the groups can be paired as the right part and the left part of the brain. The results show that the clustering based on our model successfully group the brain nodes without knowledge of the brain region.
\begin{figure}[H]
  \label{figure:elbow}
  \begin{center}
    \includegraphics[width=0.48\textwidth]{elbowmethod.png}
  \end{center}
  \caption{The elbow plot for the number of clusters}
\end{figure}

\begin{table}[H]
  \label{table:clustering}
  \center
  \begin{small}
    \footnotesize
  \begin{sc}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
Group       & \multicolumn{3}{c|}{\#1}                                                  & \multicolumn{3}{c|}{\#2}                                      \\ \hline
Brain nodes & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}"RMF\_L","Fpole\_L","Insula\_L",\\ "SupF\_L","CaudMF\_L",\\ "parsTriangularis\_L","parsOpercularis\_L",\\ "preCentral\_L","Tpole\_L","SupT\_L",\\ "SupT\_L","postCentral\_L", "SupP\_L",\\ "IP\_L","LO\_L","MOF\_L","SupF\_L",\\ "IsthmusC\_L","Precuneus\_L",\\ "Cuneus\_L","paraHippo\_L",\\ "Lingual\_L","SupT\_L","LO\_L"\end{tabular}} & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}"RMF\_R","Fpole\_R","Insula\_R",\\ "SupF\_R", "CaudMF\_R",\\ "parsTriangularis\_R" ,"parsOpercularis\_R",\\ "preCentral\_R", "Tpole\_R","SupT\_R",\\ "SupT\_R","postCentral\_R","SupP\_R",\\ "IP\_R","LO\_R","MOF\_R", "SupF\_R",\\ "IsthmusC\_R","Precuneus\_R",\\ "Cuneus\_R","paraHippo\_R",\\ "Lingual\_R","SupT\_R","LO\_R"\end{tabular}} \\ \hline
Group       & \#3    & \#4    & \#5   & \#6  & \#7 & \#8
\\ \hline
Brain nodes & \begin{tabular}[c]{@{}c@{}}"SupraM\_L" \\ "SupraM\_L" \\ "SupraM\_L" \\ "SupraM\_L"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}"SupraM\_R" \\ "SupraM\_R" \\ "SupraM\_R" \\ "SupraM\_R"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}"IT\_L" \\ "IT\_L" \\ "IT\_L"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}"IT\_R" \\ "IT\_R" \\ "IT\_R"\end{tabular}                                                          & \begin{tabular}[c]{@{}c@{}}"MT\_L" \\ "MT\_L" \\ "MT\_L"\end{tabular}                                                         & \begin{tabular}[c]{@{}c@{}}"MT\_R" \\ "MT\_R"\\ "MT\_R"\end{tabular}                                                         \\ \hline
\end{tabular}
\end{sc}
\end{small}
\caption{Result of the brain nodes clustering: the last alphabet in each node name means the left side for `L' and the right side for `R' in the brain}
\end{table}

>>>>

\subsection{Clustering result on HCP}
We perform $K$-means method to cluster brain nodes based on the suggested above method. We have the estimated latent parameter $\hat\Theta$ with the rank $(23,23,8)$ which minimizes the BIC value. The number of clusters is chosen to be eight from the elbow method in figure ~\ref{figure:elbow}.
The table in \ref{table:clustering} shows eight cluster groups and their entries. Most of brain nodes fall into the cluster $\#1$ and the cluster $\#2$ which can be represented as the left side brain and the right side brain. From the cluster $\#3$ to the cluster $\#8$ , entries of each cluster share the same name encoded in the data and play a similar role in the brain. For example, the cluster $\#3$ and the cluster $\#4$ are the left and the right side of the regions called Supramarginal gyrus. Those regions are known to be in charge of interpreting tactile sensory data and involved in perception of space and limbs location \cite{carlson2012physiology,reed1994nature}. The figure \ref{figure:brain image} shows the nodes in the dataset on the brain image. Each node is connected by edges within the same cluster and the groups can be paired as the right part and the left part of the brain. The results show that the clustering based on our model successfully group the brain nodes without knowledge of the brain region.
\begin{figure}[H]
  \label{figure:elbow}
  \begin{center}
    \includegraphics[width=0.48\textwidth]{elbowmethod.png}
  \end{center}
  \caption{The elbow plot for the number of clusters}
\end{figure}

\begin{table}[H]
  \label{table:clustering}
  \center
  \begin{small}
    \footnotesize
  \begin{sc}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
Group       & \multicolumn{3}{c|}{\#1}                                                  & \multicolumn{3}{c|}{\#2}                                      \\ \hline
Brain nodes & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}} ``RMF\_L'',``'Fpole\_L",``Insula\_L",\\ ``SupF\_L",``CaudMF\_L",\\ ``parsTriangularis\_L",``parsOpercularis\_L",\\ ``preCentral\_L",``Tpole\_L",``SupT\_L",\\ ``SupT\_L",``postCentral\_L", ``SupP\_L",\\ ``IP\_L",``LO\_L",``MOF\_L",``SupF\_L",\\ ``IsthmusC\_L",``Precuneus\_L",\\ ``Cuneus\_L",``paraHippo\_L",\\ ``Lingual\_L",``SupT\_L",``LO\_L"\end{tabular}} & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}``RMF\_R",``Fpole\_R",``Insula\_R",\\ ``SupF\_R", ``CaudMF\_R",\\ ``parsTriangularis\_R" ,``parsOpercularis\_R",\\ ``preCentral\_R", ``Tpole\_R",``SupT\_R",\\ ``SupT\_R",``postCentral\_R",``SupP\_R",\\ ``IP\_R",``LO\_R",``MOF\_R", ``SupF\_R",\\ ``IsthmusC\_R",``Precuneus\_R",\\ ``Cuneus\_R",``paraHippo\_R",\\ ``Lingual\_R",``SupT\_R",``LO\_R"\end{tabular}} \\ \hline
Group       & \#3    & \#4    & \#5   & \#6  & \#7 & \#8
\\ \hline
Brain nodes & \begin{tabular}[c]{@{}c@{}}``SupraM\_L" \\ ``SupraM\_L" \\ ``SupraM\_L" \\ ``SupraM\_L"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}``SupraM\_R" \\ ``SupraM\_R" \\ ``SupraM\_R" \\ ``SupraM\_R"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}``IT\_L" \\ ``IT\_L" \\ ``IT\_L"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}``IT\_R" \\ ``IT\_R" \\ ``IT\_R"\end{tabular}                                                          & \begin{tabular}[c]{@{}c@{}}``MT\_L" \\ ``MT\_L" \\ ``MT\_L"\end{tabular}                                                         & \begin{tabular}[c]{@{}c@{}}``MT\_R" \\ ``MT\_R"\\ ``MT\_R"\end{tabular}                                                         \\ \hline
\end{tabular}
\end{sc}
\end{small}
\caption{Result of the brain nodes clustering: the last alphabet in each node name means the left side for `L' and the right side for `R' in the brain}
\end{table}

### 2020-02-19 old_files/old_preprint/miaoyan/ordinalT_supp_v3.tex
clean up

Let $d_{\text{total}}=\prod_{k\in[K]}d_k$, and $\gamma\in[0,1]$ be a constant to be specified later.  Our strategy is to construct a finite set of tensors $\tX=\{\Theta_i\colon i=1,\ldots \}\subset \tP$ satisfying the properties of (i)-(iv) in Lemma~\ref{lem:construction}. By Lemma~\ref{lem:construction}, such a subset of tensors exist. For any tensor  $\Theta\in\tX$, let $\mathbb{P}_{\Theta}$ denote the distribution of $\tY|\Theta$, where $\tY$ is the ordinal tensor. In particular, $\mathbb{P}_{\mathbf{0}}$ is the distribution of $\tY$ induced by the zero parameter tensor $\mathbf{0}$, i.e., the distribution of $\tY$ conditional on the parameter tensor $\Theta=\mathbf{0}$. Based on the Remark for Lemma~\ref{lem:KL}, we have
\begin{equation}\label{eq:KLbound1}
KL(\mathbb{P}_{\Theta}|| \mathbb{P}_{\mathbf{0}})\leq C \FnormSize{}{\Theta}^2,
\end{equation}
where $C={(4L-6) \dot{f}^2(0)\over  A_\alpha}>0$ is a constant independent of the tensor dimension and rank.
Combining the inequality~\eqref{eq:KLbound1} with property (iii) of $\tX$, we have
\begin{equation}\label{eq:KLbound}
\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\mathbf{0}})\leq \gamma^2 R_{\max} d_{\max}.
\end{equation}
From~\eqref{eq:KLbound} and the property (i), we deduce that the condition
\begin{equation}\label{eq:totalKL}
{1\over \text{Card}(\tX)-1}\sum_{\Theta \in\tX}\text{KL}(\mathbb{P}_{\Theta}, \mathbb{P}_{\mathbf{0}})\leq \varepsilon \log_2\left\{\text{Card}(\tX)-1 \right\}
\end{equation}
holds for any $ \varepsilon \geq 0$ when $\gamma\in[0,1]$ is chosen to be sufficiently small depending on $\varepsilon$, e.g., $\gamma \leq \sqrt{3\varepsilon}$. By applying Lemma~\ref{lem:Tsybakov} to~\eqref{eq:totalKL}, and in view of the property (iv), we obtain that
\begin{equation}\label{eq:final}
\inf_{\hat \Theta}\sup_{\trueT\in \tX}\mathbb{P}\left(\FnormSize{}{\hat \Theta- \trueT}\geq  {\gamma\over 8} \min\left\{ \alpha\sqrt{d_{\text{total}}}, C^{-1/2}\sqrt{ R_{\max}d_{\max}}\right\} \right)\geq {1\over 2}\left(1-2\varepsilon-\sqrt{16 \varepsilon \over R_{\max}d_{\max}}\right).
\end{equation}
Note that $\text{Loss}(\hat \Theta, \trueT)=\FnormSize{}{\hat \Theta- \trueT}^2/d_{\text{total}}$ and $\tX\subset \tP$. By taking $\varepsilon=1/8$ and $\gamma=1/2$, we conclude from~\eqref{eq:final} that
\begin{equation}\label{eq:prob}
\inf_{\hat \Theta}\sup_{\trueT\in \tP}\mathbb{P}\left(\text{Loss}(\hat \Theta, \trueT)\geq {1 \over 256}\min\left \{ \alpha^2,  {C^{-1}R_{\max}d_{\max}\over d_{\text{total}}}\right \}\right)\geq {1\over 2}\left({3\over 4}- {2\over R_{\max}d_{\max}} \right)\geq {1\over 8}.
\end{equation}
This completes the proof.
\end{proof}

@@

\subsection{Convexity of the log-likelihood function}
\begin{thm}\label{thm:convexity}
Define the function
\begin{align}\label{eq:function}
 \logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \left\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(\theta_\omega+b_\ell)-  f(\theta_\omega+b_{\ell-1})\big]\right\},
 \end{align}
where $f(\cdot)$ satisfies Assumption~\ref{ass:link}. Then, $\logl(\Theta, \mb)$ is concave in $(\Theta,\mb)$. 
\end{thm}

\begin{proof}
Define $d_{\text{total}}=\prod_k d_k$. With a little abuse of notation, we use $(\Theta,\mb)$ to denote the length-$(d_{\text{total}}+L-1)$-vector collecting all parameters together. Let us denote a bivariate function
\begin{align}
\lambda: &\mathbb{R}^2\mapsto \mathbb{R}\\
(u,v)&\mapsto\lambda(u,v) = \log \big[f(u)-  f(v)\big].
\end{align}
It suffices to show that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$.

Suppose that the claim holds (which we will prove in the next paragraph). Based on~\eqref{eq:function}, $u,v$ are both linear functions of $(\Theta,\mb)$:
\[
u = \ma_1^T(\Theta,\mb), \quad v = \ma_2^T(\Theta,\mb), \quad \text{ for some}\ \ma_1,\ma_2\in \mathbb{R}^{d_{\text{total}}+L-1}.
\]
Then, $\lambda(u,v) = \lambda(\ma_1^T(\Theta,\mb),\ \ma_2^T(\Theta,\mb))$ is concave in $(\Theta,\mb)$ by the definition of concavity. Therefore, we can conclude that $\logl(\Theta, \mb) $ is concave in $(\Theta,\mb)$ because $\logl(\Theta, \mb)$ is the sum of $\lambda(u,v)$.

Now, we prove the concavity of $\lambda(u,v)$. Note that
\begin{equation}
\lambda(u,v) = \log\big[f(u)-f(v)\big]=\log\big[\int\mathds{1}_{[u,v]}(x)f'(x)dx\big],
\end{equation}
where $\mathds{1}_{[u,v]}$ is an indicator function that equals 1 in the interval $[u,v]$, and 0 elsewhere. Furthermore, $\mathds{1}_{[u,v]}(x)$ is log-concave in $(u,v,x)$, and by Assumption~\ref{ass:link}, $f'(x)$ is log-concave in $x$. It follows that $\mathds{1}_{(u,v)}(x)f'(x)$
is a log-concave in $(u,v,x)$. By Lemma~\ref{lem:lossconvexity}, we conclude that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$. 
\end{proof}

@@

\subsection{Auxiliary lemmas}
\label{sec:lemma}
This section collects lemmas that are useful for the proofs of the main theorems.

@@

\begin{rmk} It may seem counter-intuitive that the MSE of $\hat \Theta$ shows no dependence of $\hat \mb$ (and vice versa). In fact, from the proof of Theorem~\ref{thm:ratejoint}, the uniform constants $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}$ can be replaced by
\[
U_{\alpha,\hat \mb}=\max_{|\theta|\leq \alpha}\max_{\ell\in[L-1]}{|\dot{g}_\ell(\theta)|\over g_\ell(\theta)},\quad \text{and}\quad L_{\alpha,\hat \mb}=\min_{|\theta|\leq \alpha}\min_{\ell\in[L-1]} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}.
\]
Therefore, the accuracy of $\hat \Theta$ depends on the accuracy of $\hat \mb$ implicitly through $U_{\alpha, \hat \mb}$ and $ L_{\alpha,\hat \mb}$.

@@

\fixme{miaoyan}{Please add a few sentence to describe Figure S}
\begin{figure}[H]
  \begin{center}
    \includegraphics[width=.8\textwidth]{panel3.pdf}
  \end{center}
  \caption{Performance comparison for predicting the median label. (a,c) Prediction errors versus sample complexity $\rho=|\Omega|/d^k$ when $L=5$. (b,d) Prediction errors versus the number of ordinal levels $L$, when $\rho=0.8.$}
\end{figure}

\section{Additional results on HCP analysis}
\subsection{Clustering based on Tucker representation}\label{sec:clustering}
We perform clustering analyses based on the Tucker representation of the estimated tensor parameter $\hat\Theta$. The procedure is motivated from the higher-order extension of Principal Component Analysis (PCA) or Singular Value Decomposition (SVD). Recall that, in the matrix case, we usually perform clustering on an $m\times n$ (normalized) matrix $X$ based on the following procedure. First, we factorize $X$ into
\begin{equation}
X = U\Sigma V^T,
\end{equation}
where $\Sigma$ is a diagonal matrix and $ U,V$ are factor matrices with orthogonal columns. Second, we take each column of $V$ as a principal axis and each row in $U\Sigma$ as principal component. A subsequent multivariate clustering method (such as $K$-means) is then applied to the $m$ rows of $U\Sigma$.

We apply a similar clustering procedure to the estimated parameter tensor $\hat\Theta$. Based on Tucker representation of $\hat \Theta$, we have
\begin{equation}\label{eq:Tuckerest}
\hat \Theta = \hat \tC\times_1\hat \mM_1\times_2\cdots\times_K\hat \mM_K,
\end{equation}
where $\hat \tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is the estimated core tensor, $\hat \mM_k\in\mathbb{R}^{d_k\times r_k}$ are estimated factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\citep{kolda2009tensor}. The mode-$k$ matricization of~\eqref{eq:Tuckerest} gives
\begin{equation}
\hat \Theta_{(k)} = \hat \mM_k\hat \tC_{(k)}\left(\hat \mM_K\otimes\cdots\otimes\hat \mM_1\right),
\end{equation}
where $\hat \Theta_{(k)} , \hat \tC_{(k)}$ denote the mode-$k$ unfolding of $\hat \Theta$ and $\hat \tC$, respectively. Then, the mode-$k$ clustering can be performed as follows. First, we take colums in $\left(\hat \mM_K\otimes\cdots\otimes\hat \mM_1\right)$ as principal axes and rows in $\hat \mM_k\hat \tC_{(k)}$ as principal components. Then, we perform $K$-means clustering method to the $d_k$ rows of the matrix $\hat \mM_k\hat \tC_{(k)}$.

\subsection{Clustering results on HCP}
We perform clustering analysis on the 68 brain nodes using the the produce described in Section~\ref{sec:clustering}. Recall that our ordinal tensor method outputs the estimated parameter tensor $\hat\Theta\in\mathbb{R}^{68\times 68\times136}$ with rank $(23,23,8)$. 
We apply $K$-means to the mode-1 principal component matrix of size $68\times 184$ ($184=23\times8$). The elbow method suggests eight clusters among the 68 nodes (see Figure~\ref{figure:elbow}). Table~\ref{table:clustering} shows the eight clusters and their compositions. 
\begin{figure}[H]
  \label{figure:elbow}
  \begin{center}
    \includegraphics[width=0.48\textwidth]{elbowmethod.png}
  \end{center}
  \caption{The elbow plot for the number of clusters}
\end{figure}

>>>>

Let $d_{\text{total}}=\prod_{k\in[K]}d_k$, and $\gamma\in[0,1]$ be a constant to be specified later.  Our strategy is to construct a finite set of tensors $\tX=\{\Theta_i\colon i=1,\ldots \}\subset \tP$ satisfying the properties of (i)-(iv) in Lemma~\ref{lem:construction}. By Lemma~\ref{lem:construction}, such a subset of tensors exist. For any tensor  $\Theta\in\tX$, let $\mathbb{P}_{\Theta}$ denote the distribution of $\tY|\Theta$, where $\tY$ is the ordinal tensor. In particular, $\mathbb{P}_{\mathbf{0}}$ is the distribution of $\tY$ induced by the zero parameter tensor $\mathbf{0}$, i.e., the distribution of $\tY$ conditional on the parameter tensor $\Theta=\mathbf{0}$. Based on the Remark for Lemma~\ref{lem:KL}, we have
\begin{equation}\label{eq:KLbound1}
KL(\mathbb{P}_{\Theta}|| \mathbb{P}_{\mathbf{0}})\leq C \FnormSize{}{\Theta}^2,
\end{equation}
where $C={(4L-6) \dot{f}^2(0)\over  A_\alpha}>0$ is a constant independent of the tensor dimension and rank.
Combining the inequality~\eqref{eq:KLbound1} with property (iii) of $\tX$, we have
\begin{equation}\label{eq:KLbound}
\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\mathbf{0}})\leq \gamma^2 R_{\max} d_{\max}.
\end{equation}
From~\eqref{eq:KLbound} and the property (i), we deduce that the condition
\begin{equation}\label{eq:totalKL}
{1\over \text{Card}(\tX)-1}\sum_{\Theta \in\tX}\text{KL}(\mathbb{P}_{\Theta}, \mathbb{P}_{\mathbf{0}})\leq \varepsilon \log_2\left\{\text{Card}(\tX)-1 \right\}
\end{equation}
holds for any $ \varepsilon \geq 0$ when $\gamma\in[0,1]$ is chosen to be sufficiently small depending on $\varepsilon$, e.g., $\gamma \leq \sqrt{3\varepsilon}$. By applying Lemma~\ref{lem:Tsybakov} to~\eqref{eq:totalKL}, and in view of the property (iv), we obtain that
\begin{equation}\label{eq:final}
\inf_{\hat \Theta}\sup_{\trueT\in \tX}\mathbb{P}\left(\FnormSize{}{\hat \Theta- \trueT}\geq  {\gamma\over 8} \min\left\{ \alpha\sqrt{d_{\text{total}}}, C^{-1/2}\sqrt{ R_{\max}d_{\max}}\right\} \right)\geq {1\over 2}\left(1-2\varepsilon-\sqrt{16 \varepsilon \over R_{\max}d_{\max}}\right).
\end{equation}
Note that $\text{Loss}(\hat \Theta, \trueT)=\FnormSize{}{\hat \Theta- \trueT}^2/d_{\text{total}}$ and $\tX\subset \tP$. By taking $\varepsilon=1/8$ and $\gamma=1/2$, we conclude from~\eqref{eq:final} that
\begin{equation*}\label{eq:prob}
\inf_{\hat \Theta}\sup_{\trueT\in \tP}\mathbb{P}\left(\text{Loss}(\hat \Theta, \trueT)\geq {1 \over 256}\min\left \{ \alpha^2,  {C^{-1}R_{\max}d_{\max}\over d_{\text{total}}}\right \}\right)\geq {1\over 2}\left({3\over 4}- {2\over R_{\max}d_{\max}} \right)\geq {1\over 8}.
\end{equation*}
This completes the proof.
\end{proof}

@@

\subsection{Convexity of the log-likelihood function}
\begin{thm}\label{thm:convexity}
Define the function
\begin{align}\label{eq:function}
 \logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \left\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(\theta_\omega+b_\ell)-  f(\theta_\omega+b_{\ell-1})\big]\right\},
 \end{align}
where $f(\cdot)$ satisfies Assumption~\eqref{eq:ass}. Then, $\logl(\Theta, \mb)$ is concave in $(\Theta,\mb)$. 
\end{thm}

\begin{proof}
Define $d_{\text{total}}=\prod_k d_k$. With a little abuse of notation, we use $(\Theta,\mb)$ to denote a length-$(d_{\text{total}}+L-1)$-vector collecting all parameters together. Let us denote a bivariate function
\begin{align}
\lambda: &\mathbb{R}^2\mapsto \mathbb{R}\\
(u,v)&\mapsto\lambda(u,v) = \log \big[f(u)-  f(v)\big].
\end{align}
It suffices to show that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$.

Suppose that the claim holds (which we will prove in the next paragraph). Based on~\eqref{eq:function}, $u,v$ are both linear functions of $(\Theta,\mb)$:
\[
u = \ma_1^T(\Theta,\mb), \quad v = \ma_2^T(\Theta,\mb), \quad \text{ for some}\ \ma_1,\ma_2\in \mathbb{R}^{d_{\text{total}}+L-1}.
\]
Then, $\lambda(u,v) = \lambda(\ma_1^T(\Theta,\mb),\ \ma_2^T(\Theta,\mb))$ is concave in $(\Theta,\mb)$ by the definition of concavity. Furthermore, we can conclude that $\logl(\Theta, \mb) $ is concave in $(\Theta,\mb)$ because $\logl(\Theta, \mb)$ is the sum of $\lambda(u,v)$.

Now, we prove the concavity of $\lambda(u,v)$. Note that
\begin{equation*}
\lambda(u,v) = \log\big[f(u)-f(v)\big]=\log\big[\int\mathds{1}_{[u,v]}(x)f'(x)dx\big],
\end{equation*}
where $\mathds{1}_{[u,v]}$ is an indicator function that equals 1 in the interval $[u,v]$, and 0 elsewhere. Furthermore, $\mathds{1}_{[u,v]}(x)$ is log-concave in $(u,v,x)$ and $f'(x) = \frac{e^x}{(1+e^x)^2}$ is log-concave in $x$. The convexities imply that $\mathds{1}_{(u,v)}(x)f'(x)$
is a log-concave in $(u,v,x)$. By Lemma~\ref{lem:lossconvexity}, we conclude that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$. 
\end{proof}

@@

\subsection{Auxiliary lemmas}
\label{sec:lemma}

We begin with various notion of tensor norms that are useful for the proofs of the main theorems.

@@

\begin{rmk} It may seem counter-intuitive that the MSE of $\hat \Theta$ shows no dependence of $\hat \mb$ (and vice versa). In fact, from the proof of Theorem~\ref{thm:ratejoint}, the uniform constants $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}$ can be replaced by
\[
U_{\alpha,\hat \mb}=\max_{|\theta|\leq \alpha}\max_{\ell\in[L-1]}{|\dot{g}_\ell(\theta)\over g_\ell(\theta)|},\quad \text{and}\quad L_{\alpha,\hat \mb}=\min_{|\theta|\leq \alpha}\min_{\ell\in[L-1]} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}.
\]
Therefore, the accuracy of $\hat \Theta$ depends on the accuracy of $\hat \mb$ implicitly through $U_{\alpha, \hat \mb}$ and $ L_{\alpha,\hat \mb}$.

@@

\Fixme{miaoyan}{Please add a few sentence to describe Figure S}
\begin{figure}[H]
  \begin{center}
    \includegraphics[width=0.48\textwidth]{panel3.pdf}
  \end{center}
  \caption{Performance comparison in predicting the median label. (a,c) Prediction errors versus sample complexity $\rho=|\Omega|/d^k$ when $L=5$. (b,d) Prediction errors versus the number of ordinal levels $L$, when $\rho=0.8.$}
\end{figure}

\section{Additional results on HCP analysis}
\subsection{Clustering based on Tucker representation}\label{sec:clustering}
We perform clustering analyses based on the Tucker representation of the estimated tensor parameter $\hat\Theta$. The procedure is motivated from the higher-order extension of Principal Component Analysis (PCA) or Singular Value Decomposition (SVD). Recall that, in the matrix case, we usually perform clustering on an $m\times n$ (normalized) matrix $X$ based on the following procedure. First, we factorize $X$ into
\begin{equation}
X = U\Sigma V^T,
\end{equation}
where $\Sigma$ is a diagonal matrix and $ U,V$ are factor matrices with orthogonal columns. Second, we take each column of $V$ as a principal axis and $U\Sigma$ as principal components. A subsequent multivariate clustering method (such as $K$-means) is then applied to the $m$ rows of $U\Sigma$.

We apply a similar clustering procedure to the estimated parameter tensor $\hat\Theta$. Based on Tucker decomposition, we have
\begin{equation}\label{eq:Tuckerest}
\hat \Theta = \hat \tC\times_1\hat \mM_1\times_2\cdots\times_K\hat \mM_K,
\end{equation}
where $\hat \tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is the estimated core tensor, $\hat \mM_k\in\mathbb{R}^{d_k\times r_k}$ are estimated factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\citep{kolda2009tensor}. The mode-$k$ matricization of~\eqref{eq:Tuckerest} gives
\begin{equation}
\hat \Theta_{(k)} = \hat \mM_k\hat \tC_{(k)}\left(\hat \mM_K\otimes\cdots\otimes\hat \mM_1\right),
\end{equation}
where $\hat \Theta_{(k)} , \hat \tC_{(k)}$ denote the mode-$k$ unfolding of $\hat \Theta$ and $\hat \tC$, respectively. Then, the mode-$k$ clustering can be performed as follows. First, we take $\left(\hat \mM_K\otimes\cdots\otimes\hat \mM_1\right)$ as principal axes and $\hat \mM_k\hat \tC_{(k)}$ as principal components. Then, we perform $K$-means clustering method to the $d_k$ rows of the principal components $\hat \mM_k\hat \tC_{(k)}$.

\subsection{Clustering results on HCP}
We perform clustering analysis to the 68 brain nodes using the the produce described in Section~\ref{sec:clustering}. Recall that our ordinal tensor method outputs the estimated parameter tensor $\hat\Theta\in\mathbb{R}^{68\times 68\times136}$ with rank $(23,23,8)$. 
The number of clusters in the $K$-means is chosen based on the elbow method (see figure~\ref{figure:elbow}). Table~\ref{table:clustering} shows the eight clusters of brain nodes. 
\begin{figure}[H]
  \label{figure:elbow}
  \begin{center}
    \includegraphics[width=0.48\textwidth]{elbowmethod.png}
  \end{center}
  \caption{The elbow plot for the number of clusters}
\end{figure}

### 2020-02-19 old_files/old_preprint/old_arXiv/ordinalTv4.tex
clean up

\usepackage{xr}
\externaldocument{ordinalT_supp_v3}

@@

\begin{thm}[Minimax lower bound]\label{thm:minimax}
Assume the same set-up as in Theorem~\ref{thm:rate}, and $d_{\max}=\max_k d_k \geq 8$. Let $\inf_{\hat \Theta}$ denote the infimum over all estimators $\hat \Theta\in\tP$ based on the ordinal tensor observation $\tY\in[L]^{d_1\times \cdots \times d_K}$. Then, under the model~\eqref{eq:model},
\begin{align}\label{eq:lower}
\inf_{\hat \Theta }\sup_{\trueT\in\tP} \mathbb{P}\Big\{& \text{MSE}(\hat \Theta, \trueT) \\
& \geq {1\over 256} \min\left( \alpha^2, \ { Cr_{\max}d_{\max} \over \prod_k d_k} \right) \Big\} \geq {1\over 8},
\end{align}
where $C=C(\alpha, L, f,\mb)>0$ is a constant independent of tensor dimension and the rank.
\end{thm}
We see that the lower bound matches the upper bound in~\eqref{eq:rate} on the polynomial order of tensor dimension. Therefore, our estimator~\eqref{eq:estimator} is order-optimal.

@@

\vspace{-.1cm}
\subsection{Comparison with alternative methods}
\vspace{-.1cm}
Next, we compare our ordinal tensor method ({\bf Ordinal-T}) with three popular low-rank methods:

@@

\vspace{-.1cm}
\subsection{Human Connectome Project (HCP)}
\vspace{-.1cm}
Each entry in the HCP dataset takes value on a nominal scale, \{{\it high, moderate, low}\}, indicating the strength level of fiber connection. We convert the dataset to a 3-level ordinal tensor $\tY\in[3]^{68\times 68\times 136}$ and apply the ordinal tensor method with a logistic link function. The BIC suggests $\mr = (23,23,8)$ with $\logl(\hat{\Theta},\hat{\mb}) = -216,646$. Based on the estimated Tucker factors $\{\hat \mM_k\}$, we perform a clustering analysis via K-mean on the brain nodes (see detailed procedure in the Supplement). We find that the 68 brain nodes are grouped into eight clusters, and the clustering capture the spatial separation of brain nodes very well (see the Supplement). In particular, the top two clusters represent the left and right hemispheres; the smaller clusters represent local regions driving by similar nodes. For example, the cluster IV consists of brain nodes {\it postcentral}, {\it posteriorcingulate}, {\it pericalcarine} and {\it temporalpole}, all of which are located at {\it SupraM}. The identified similarity among brain nodes suggests the applicability of our method.

>>>>

@@

\begin{thm}[Minimax lower bound]\label{thm:minimax}
Assume the same set-up as in Theorem~\ref{thm:rate}, and $d_{\max}=\max_k d_k \geq 8$. Let $\inf_{\hat \Theta}$ denote the infimum over all estimators $\hat \Theta\in\tP$ based on the ordinal tensor observation $\tY\in[L]^{d_1\times \cdots \times d_K}$. Then, under the model~\eqref{eq:model},
\begin{align}\label{eq:lower}
\inf_{\hat \Theta }\sup_{\trueT\in\tP} \mathbb{P}\Big\{& \text{MSE}(\hat \Theta, \trueT) \\
& \geq c \min\left( \alpha^2, \ { Cr_{\max}d_{\max} \over \prod_k d_k} \right) \Big\} \geq {1\over 8},
\end{align}
where $C=C(\alpha, L, f,\mb)>0$ and $c>0$ are constants independent of tensor dimension and the rank.
\end{thm}
We see that the lower bound matches the upper bound in~\eqref{eq:rate} on the polynomial order of tensor dimension. Therefore, our estimator~\eqref{eq:estimator} is order-optimal.

@@

\vspace{-.1cm}
\subsection{Comparison with alternative methods}\label{sec:compare}
\vspace{-.1cm}
Next, we compare our ordinal tensor method ({\bf Ordinal-T}) with three popular low-rank methods:

@@

\vspace{-.1cm}
\subsection{Human Connectome Project (HCP)}
\vspace{-.1cm}
Each entry in the HCP dataset takes value on a nominal scale, \{{\it high, moderate, low}\}, indicating the strength level of fiber connection. We convert the dataset to a 3-level ordinal tensor $\tY\in[3]^{68\times 68\times 136}$ and apply the ordinal tensor method with a logistic link function. The BIC suggests $\mr = (23,23,8)$ with $\logl(\hat{\Theta},\hat{\mb}) = -216,646$. Based on the estimated Tucker factors $\{\hat \mM_k\}$, we perform a clustering analysis via K-mean on the brain nodes (see detailed procedure in the Supplement). We find that the 68 brain nodes are grouped into eight clusters, and the clustering capture the spatial separation of brain nodes very well (see the Supplement). In particular, the top two clusters represent the left and right hemispheres; the smaller clusters represent local regions driving by similar nodes. For example, the cluster IV consists of brain nodes all of which are called {\it R.SUPRAMARGINAL} and located at the same region. The identified similarity among brain nodes suggests the applicability of our method.
